---
title: "NCAA Project"
author: "Amber Duevel"
date: "9/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(car)
```

```{r read in file, message = FALSE, warning = FALSE}
ace_vb<-read_csv("VB_aces-1.csv")
assists_vb<-read_csv("VB_assists-1.csv")
blocks_vb<-read_csv("VB_blocks-1.csv")
digs_vb<-read_csv("VB_digs-1.csv")
hit_perc_vb<-read_csv("VB_hit_perc-1.csv")
kills_vb<-read_csv("VB_kills-1.csv")
opp_hits_vb<-read_csv("VB_opp_hits-1.csv")
WL_perc_vb<-read_csv("VB_WL_perc-1.csv")
```

```{r join data}
NCAA_full_set <- ace_vb %>% right_join(assists_vb, by="Team") %>% 
  right_join(blocks_vb, by="Team") %>% 
  right_join(digs_vb, by="Team") %>% 
  right_join(hit_perc_vb, by="Team") %>% 
  right_join(kills_vb, by="Team") %>% 
  right_join(opp_hits_vb, by="Team") %>% 
  right_join(WL_perc_vb, by="Team")
```

```{r clean data}
names(NCAA_full_set)
NCAA_full_set <- NCAA_full_set %>% select(-c("...1.x", "...1.y", "S.y", "...1.x.x","S.x.x", "...1.y.y", "S.y.y", "...1.x.x.x", "S.x.x.x","...1.y.y.y", "S.y.y.y", "...1.x.x.x.x", "S","...1.y.y.y.y", "Rank.x", "Rank.y", "Rank.x.x", "Rank.y.y", "Rank.x.x.x", "Rank.y.y.y", "Rank.x.x.x.x", "Rank.y.y.y.y", "Kills.y")) 
NCAA_full_set = distinct(NCAA_full_set, Team, .keep_all = TRUE)
```


```{r rename}
NCAA_full_set <- rename(NCAA_full_set,c("Team_Name"="Team",
                                        "Sets"="S.x",
                                        "Aces"="Aces",
                                        "Per_Set_x"="Per Set.x",
                                        "Assists"="Assists", 
                                        "Per_Set_y"="Per Set.y",
                                        "Block_Solos"="Block Solos",
                                        "Block_Assists"="Block Assists",
                                        "Triple_Block"="TB",
                                        "Digs"="Digs",
                                        "Per_Set_xx"="Per Set.x.x",
                                        "Kills"="Kills.x",
                                        "Errors"="Errors", 
                                        "Total_Attacks"="Total Attacks",
                                        "Hitting_Percentage"="Pct..x",
                                        "Percentage_Set"="Per Set.y.y",
                                        "Opponents_Kill"="Opp Kills",
                                        "Opponents_Errors"="Opp Errors",
                                        "Opponents_Attacks"="Opp Attacks",
                                        "Opponents_Percentage"="Opp Pct",
                                        "Win"="W",
                                        "Lost"="L",
                                        "Win_Percentage"="Pct..y"))
```
Intro:

In this project we focused on one main research question; what team statistic(s) is most highly predictive of team’s wins and losses. We answered this question by testing for the R2 score between win percentages and each team statistic variable. The team statistic variables are things like aces, blocks, assists, etc. This way we can see if any of the variables have a correlation with win percentage. 

```{r ace}
ace_lm = lm((Win_Percentage+0.01)~Aces, data = NCAA_full_set)
summary(ace_lm)

ace_plot = ggplot(aes(x = Aces, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
ace_plot

par(mfrow = c(2,2))
plot(ace_lm)
boxCox(ace_lm)

# suggests log transformation
ace_trans = lm(sqrt(Win_Percentage)~Aces, data = NCAA_full_set)
summary(ace_trans)

ace2_plot = ggplot(aes(x = Aces, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
ace2_plot

par(mfrow = c(2,2))
plot(ace_trans)
```


```{r assists}
assists_lm = lm(Win_Percentage+0.01~Assists, data = NCAA_full_set)
summary(assists_lm)

assists_plot = ggplot(aes(x = Assists, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
assists_plot

par(mfrow = c(2,2))
plot(assists_lm)
boxCox(assists_lm)

# I guess it suggests log transformation. Includes 0.75, but not 0.5 or 1. 
assists_trans = lm(log(Win_Percentage + 0.01)~Assists, data = NCAA_full_set)
summary(assists_trans)

assists2_plot = ggplot(aes(x = Assists, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
assists2_plot

par(mfrow = c(2,2))
plot(assists_trans)
```

```{r block solo}
block_solo_lm = lm(Win_Percentage+0.01~Block_Solos, data = NCAA_full_set)
summary(block_solo_lm)

block_solo_plot = ggplot(aes(x = Block_Solos, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
block_solo_plot

par(mfrow = c(2,2))
plot(block_solo_lm)
boxCox(block_solo_lm)

# suggests log transformation
block_solo_trans = lm(log(Win_Percentage + 0.01)~Block_Solos, data = NCAA_full_set)
summary(block_solo_trans)

block_solo2_plot = ggplot(aes(x = Block_Solos, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
block_solo2_plot

par(mfrow = c(2,2))
plot(block_solo_trans)
```

```{r block assists}
block_assists_lm = lm(Win_Percentage+0.01~Block_Assists, data = NCAA_full_set)
summary(block_assists_lm)

block_assists_plot = ggplot(aes(x = Block_Assists, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
block_assists_plot

par(mfrow = c(2,2))
plot(block_assists_lm)
boxCox(block_assists_lm)

# suggests log transformation
block_assists_trans = lm(log(Win_Percentage + 0.01)~Block_Assists, data = NCAA_full_set)
summary(block_assists_trans)

block_assists2_plot = ggplot(aes(x = Block_Assists, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
block_assists2_plot

par(mfrow = c(2,2))
plot(block_assists_trans)
```

```{r triple block}
triple_block_lm = lm(Win_Percentage+0.01~Triple_Block, data = NCAA_full_set)
summary(triple_block_lm)

triple_block_plot = ggplot(aes(x = Triple_Block, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
triple_block_plot

par(mfrow = c(2,2))
plot(triple_block_lm)
boxCox(triple_block_lm)

# suggests log transformation
triple_block_trans = lm(log(Win_Percentage + 0.01)~Triple_Block, data = NCAA_full_set)
summary(triple_block_trans)

triple_block2_plot = ggplot(aes(x = Triple_Block, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
triple_block2_plot

par(mfrow = c(2,2))
plot(triple_block_trans)
```

```{r digs}
digs_lm = lm(Win_Percentage+0.01~Digs, data = NCAA_full_set)
summary(digs_lm)

digs_plot = ggplot(aes(x = Digs, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
digs_plot

par(mfrow = c(2,2))
plot(digs_lm)
boxCox(digs_lm)

# suggests log transformation
digs_trans = lm(log(Win_Percentage + 0.01)~Digs, data = NCAA_full_set)
summary(digs_trans)

digs2_plot = ggplot(aes(x = Digs, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
digs2_plot

par(mfrow = c(2,2))
plot(digs_trans)
```

```{r kills}
kills_lm = lm(Win_Percentage+0.01~Kills, data = NCAA_full_set)
summary(kills_lm)

kills_plot = ggplot(aes(x = Kills, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
kills_plot

par(mfrow = c(2,2))
plot(kills_lm)
boxCox(kills_lm)

# suggests log transformation
kills_trans = lm(log(Win_Percentage + 0.01)~Kills, data = NCAA_full_set)
summary(kills_trans)

kills2_plot = ggplot(aes(x = Kills, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
kills2_plot

par(mfrow = c(2,2))
plot(kills_trans)
```

```{r errors}
errors_lm = lm(Win_Percentage+0.01~Errors, data = NCAA_full_set)
summary(errors_lm)

errors_plot = ggplot(aes(x = Errors, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
errors_plot

par(mfrow = c(2,2))
plot(errors_lm)
boxCox(errors_lm)

# suggests log transformation
errors_trans = lm(log(Win_Percentage + 0.01)~Errors, data = NCAA_full_set)
summary(errors_trans)

errors2_plot = ggplot(aes(x = Errors, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
errors2_plot

par(mfrow = c(2,2))
plot(errors_trans)
```

```{r total attacks}
total_attacks_lm = lm(Win_Percentage+0.01~Total_Attacks, data = NCAA_full_set)
summary(total_attacks_lm)

total_attacks_plot = ggplot(aes(x = Total_Attacks, y = Win_Percentage), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
total_attacks_plot

par(mfrow = c(2,2))
plot(total_attacks_lm)
boxCox(total_attacks_lm)

# suggests log transformation
total_attacks_trans = lm(log(Win_Percentage + 0.01)~Total_Attacks, data = NCAA_full_set)
summary(total_attacks_trans)

total_attacks2_plot = ggplot(aes(x = Total_Attacks, y = log(Win_Percentage)), data = NCAA_full_set) +
  geom_point() +
  geom_smooth(method = "lm")
total_attacks2_plot

par(mfrow = c(2,2))
plot(total_attacks_trans)
```
